---
title: "Calculate Selection Index"
author: "Chris Simoes"
date: "3/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(results="hold")
knitr::opts_chunk$set(collapse=TRUE)
knitr::opts_chunk$set(fig.show="hold")
knitr::opts_chunk$set(tidy.opts = list(wrap=FALSE))
```


## Selecion Index

This is a script to calculate the best selection index when using more than one trait. 
It is necessary to know the heritability, correlation and the economic value for each trait.

###Script setup
Install packages, start the README.md, set the random seed  
```{r Script setup, echo=FALSE, include=FALSE}
ip <- installed.packages()
packages_used <- c("tidyverse", "devtools", "here")

ip <- installed.packages()
for (package in packages_used){
  if (!(package %in% ip[,"Package"])) install.packages(package)
}#END packages_used

here::i_am("analysis/SelectionIndex.Rmd")

source(here::here("code/addToREADME.R"))

addToREADME(paste0("## ", rmarkdown::metadata$title), append=F)
addToREADME(c(date(), ""))

packages_info <- ip[packages_used, c("Package", "Version", "Built")]
addToREADME(c("The packages used in this script are:", "Package, Version, Built"))
apply(packages_info, 1, function(vec) addToREADME(paste(vec, collapse=" ")))
addToREADME("")

random_seed <- 835246
set.seed(random_seed)
addToREADME(paste("The random seed is", random_seed))
```
## Defining paramenters

You can also embed plots, for example:

```{r parameters, echo=TRUE}

nTraits = 4
VarPheno <- c(6.37, 0.28,3.24, 2.64)
ecoValue <- c(1,1,1,0.5)
H2 <- c(0.36,0.44,0.53,0.19)
# The number of terms in rA and rP must be:
# n(n-1)/2
# Where n is the number of traits
dimension <- nTraits*(nTraits-1)/2
cat("Regression coefficeints must have ",dimension, "elemnts.")
rA <- c(0.29, 0.32, 0.41, 0.36, 0.29,0.21)
rP <- c(0.45, 0.48,0.54,0.39,0.28,0.51)

addToREADME(c("", "Main Parameters",
              paste("Traits", nTraits),
              paste("Heritability", H2),
              paste("Pheno Variance", VarPheno),
              paste("Economic Value", ecoValue),
              paste("Additive Coefficiente", rA),
              paste("Pheno coefficient", rP))
            )


```

## Calculating Index

This part is using the equation to calculate the index. It has three items:
1 - Phenotypic matrix (with variance in the diagonal and the correlation between traits in off diagonal).
2 - Genotypic matrix (it is a matrix with one col and nTraits lines). Calculate the additive variance using heritability formula.
3 - Economic value is not matrix (They are scalar values to muliply each index).

```{r selection index, echo=TRUE}

# 1 Phenotypic matrix
#The of diagonal is defined by the covariation between the traits.
# cov(1,2) = rP x sqrt(varPheno[1] x varPheno[2])
PhenoMatrix <- matrix(data = NA, nrow = nTraits, ncol = nTraits)

for (i in 1:nTraits){
  if (i < nTraits){
    for (j in (i+1):nTraits){
      offDiagonal <- rP[i]*(sqrt(VarPheno[i] * VarPheno[j]))
      PhenoMatrix[i,j] <- offDiagonal
      PhenoMatrix[j,i] <- offDiagonal
    }
  }
  PhenoMatrix[i,i] <- VarPheno[i]
}


InvPhenoMatrix <- solve(PhenoMatrix) #solve is a command to calculate the inverted matrix

# 2 Genotypic matrix
GenoMatrix <- matrix(data=NA, ncol = 1, nrow = nTraits)

for (i in 1:nTraits){
   GenoMatrix[i,1] <- H2[i]*VarPheno[i]

}

# 3 Economic value
# Already defined as ecoValue

# Calculate the Index
SelectionIndex <- InvPhenoMatrix %*% GenoMatrix

# Multiplying the economic value

for (i in 1:nTraits) {
  SelectionIndex[i,1] <- SelectionIndex[i,1]*ecoValue[i]
}


print(SelectionIndex)

addToREADME(c("Chunk Initiate simulation",
  paste("Index for trait 1", paste(SelectionIndex[1,1], collapse=" ")),
  paste("Index for trait 2", paste(SelectionIndex[2,1], collapse=" "))
))


```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
